%% = https://cs.brown.edu/courses/csci1490/slides/CS149-BranchBound.pdf

%% - Search Strategies

%===========================================================================%
\begin{frame}
\frametitle{Search Strategies}

– Depth First Search
• Finds feasible solutions quickly.
• Is very memory efficient.
• Can easily get stuck in sub-optimal parts of the
search space.
\end{frame}
%===========================================================================%
\begin{frame}
\frametitle{Search Strategies}
– Best First Search
• Look at the node with best relaxation value next.
• Is provably optimal in the sense that it never visits a
node that could be pruned otherwise.
• A lot of jumping is necessary and memory
requirements are prohibitively large (often search
degenerates to breadth first search).
\end{frame}
%===========================================================================%
\begin{frame}
\frametitle{Search Strategies}
– Depth First Search with Best Backtracking
• Is a mix of both depth and best first search: perform depth first
search until a leaf is found, then backtrack to the node with best
relaxation value and so on.
• Much less jumping than best first search.
• Is more memory efficient than best first search, but less than
DFS – could still be very memory intensive.
\end{frame}
%===========================================================================%
\begin{frame}
\frametitle{Search Strategies}

– Least Discrepancy Search
• Follow DFS with heuristic branching direction selection.
Investigate leaves in order of increasing discrepancy wrt that
heuristic.
• Memory requirements are within limits.
• Often finds good solutions early in the search.
